Supplementary information

### 2018-06-12 supplementary_information.Rmd
Setup drake. Work on supplementary materials.

---
title: Estimating trait means and covariances through multiple imputation
author:
  - >
    Alexey N. Shiklomanov,
    Elizabeth M. Cowdery,
    Michael Bahn,
    Chaeho Byun,
    Joseph Craine,
    Andrés Gonzalez-Melo,
    Steven Jansen,
    Nathan Kraft,
    Koen Kramer,
    Vanessa Minden,
    Ülo Niinemets,
    Yusuke Onoda,
    Enio Egon Sosinski,
    Nadejda A. Soudzilovskaia,
    Michael C. Dietze
output: pdf_document
---

# Introduction

Under _single_ imputation, the imputation step occurs once outside of the MCMC loop.
In _multiple_ imputation, as implemented in this paper, the imputation step is part of the MCMC loop,
such that imputed data values are conditioned on the current state of the parameters and vice-versa.
In the figure below, $Y$ is the original data (with missing values), $Y^*$ is the original data with missing values imputed, $\mu$ and $\Sigma$ are the estimates of the

```{tikz diagram, echo = FALSE}
\usetikzlibrary{positioning, calc, fit}
\begin{tikzpicture}[
	every node/.style = {shape = rectangle, draw = black},
	scale = 1.1
]
\node (Y) at (-1, 3) {$Y$};
\node (data params) at (-1, 2) {$\mu_p$, $\Sigma_p$};
\node (Ystar) at (-1, 1) {$Y^*$};
\node (guess params) [align = center] at (1, 1) {Initial guess\\$\mu^*_0$ $\Sigma^*_0$};
\node (mu star) at (0, 0) {$\mu^*_t \mid Y^*, \Sigma^*_{t-1}$};
\node (sigma star) at (0, -1) {$\Sigma^*_t \mid Y^*, \mu^*_t$};
\path [->] (Y) edge (data params);
\path [->] (data params) edge (Ystar);
\path [->] (Ystar) edge (mu star);
\path [->] (guess params) edge (mu star);
\path [->] (mu star) edge (sigma star);
\path [->, out = 0, in = 0] (sigma star.east) edge
	node (t plus one) [right, draw = none, font = {\footnotesize}] {$t = t + 1$}
	(mu star.east);
\node (mcmc loop) [fit = (mu star)(sigma star)(t plus one), shape = rectangle, draw = black!50] {};
\node (mcmc loop text) [align = center, anchor = west, draw = none] at (mcmc loop.east) {MCMC\\loop};
\node (title) [anchor = north, draw = none] at (mcmc loop.south) {Single imputation};
\end{tikzpicture}
\hspace{3cm}
\begin{tikzpicture}[
	every node/.style = {shape = rectangle, draw = black},
	scale = 1.1
]
\node (Y) at (-1, 1) {$Y$};
\node (guess params) [align = center] at (1, 1) {Initial guess\\$\mu^*_0$ $\Sigma^*_0$};
\node (Ystar) at (0, 0) {$Y^*_t \mid Y, \mu^*_{t-1}, \Sigma^*_{t-1}$};
\node (mu star) at (0, -1) {$\mu^*_t \mid Y^*_t, \Sigma^*_{t-1}$};
\node (sigma star) at (0, -2) {$\Sigma^*_t \mid Y^*_t, \mu^*_t$};
\path [->] (Y) edge (Ystar);
\path [->] (guess params) edge (Ystar);
\path [->] (Ystar) edge (mu star);
\path [->] (mu star) edge (sigma star);
\path [->, out = 0, in = 0] (sigma star.east) edge
	node (t plus one) [right, draw = none, font = {\footnotesize}] {$t = t + 1$}
	(Ystar.east);
\node (mcmc loop) [fit = (Ystar)(mu star)(sigma star)(t plus one), shape = rectangle, draw = black!50] {};
\node (mcmc loop text) [align = center, anchor = west, draw = none] at (mcmc loop.east) {MCMC\\loop};
\node (title) [anchor = north, draw = none] at (mcmc loop.south) {Multiple imputation};
\end{tikzpicture}
```

# Demonstration

Consider two positively correlated traits, A and B.

```{r create_traits}
library(mvtraits)

true_mu <- c(0, 0)
true_cov <- matrix(c(1, 0.8, 0.8, 1), nrow = 2)
Y_all <- random_mvnorm(1000, true_mu, true_cov)
colnames(Y_all) <- c("A", "B")
plot(Y_all, pch = 19)
```

Randomly remove some values from the data.

```{r remove_na}
Y <- Y_all
Y[sample.int(700)] <- NA
plot(Y_all, pch = 19, col = "grey")
points(Y, pch = 19, col = "black")
```

Set an uninformative multivariate normal prior on $\mu$...
```{r prior_mu}
mu_0 <- c(0, 0)
Sigma_0 <- diag(10, 2)
```

...and an uninformative Wishart prior on $\Sigma$.
```{r prior_sigma}
v0 <- 2
S0 <- diag(10, 2)
```

Take an initial guess at the mean and covariance by drawing from their priors.

```{r initial}
mu_star <- random_mvnorm(1, mu_0, Sigma_0)[1,]
Sigma_star <- rWishart(1, v0, S0)[,,1]
```

Impute values based on this initial guess.

```{r impute}
Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)
plot(Y_all, pch = 19, col = "grey")
points(Ystar, pch = 19, col = "red")
points(Y, pch = 19, col = "black")
```

Clearly, these initial values are a bad fit to the data.
Nevertheless, draw $\mu$ and $\Sigma$ conditioned on this set of imputed values.

```{r draw}
mu_star <- draw_mu(Ystar, Sigma_star, mu_0, Sigma_0)
Sigma_star <- draw_Sigma(Ystar, mu_star, v0, S0)
```

Draw a new set of imputed values, conditioned on the current $\mu^*$ and $\Sigma^*$.

```{r impute_2}
Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)
plot(Y_all, pch = 19, col = "grey")
points(Ystar, pch = 19, col = "red")
points(Y, pch = 19, col = "black")
```

Doing this repeatedly (in a loop) gradually improves the estimates of the covariance.

```{r loop}
par(mfrow = c(5, 4), mar = c(3, 3, 2, 0.1), cex = 0.4)
for (i in 1:20) {
  mu_star <- draw_mu(Ystar, Sigma_star, mu_0, Sigma_0)[1,]
  Sigma_star <- draw_Sigma(Ystar, mu_star, v0, S0)
  Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)
  title <- sprintf("iteration: %d, covariance: %.2f", i, Sigma_star[2, 1])
  plot(Y_all, pch = 19, col = "grey", xlab = "", ylab = "", main = title)
  points(Ystar, pch = 19, col = "red")
  points(Y, pch = 19, col = "black")
}
```

After a lot of MCMC samples, we can generate a distribution of covariance values, which provides an uncertainty estimate on the covariance.

```{r cov_estimate}
cov_ab <- numeric(1000)
for (i in seq_along(cov_ab)) {
  mu_star <- draw_mu(Ystar, Sigma_star, mu_0, Sigma_0)[1,]
  Sigma_star <- draw_Sigma(Ystar, mu_star, v0, S0)
  Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)
  cov_ab[i] <- Sigma_star[2, 1]
}
hist(cov_ab, xlab = "Covariance estimate", ylab = "Count", main = "Covariance estimate")
abline(v = true_cov[2, 1], col = "red", lwd = 2)
```

>>>>

---
title: Supplementary information
author:
  - >
    Alexey N. Shiklomanov,
    Elizabeth M. Cowdery,
    Michael Bahn,
    Chaeho Byun,
    Joseph Craine,
    Andrés Gonzalez-Melo,
    Steven Jansen,
    Nathan Kraft,
    Koen Kramer,
    Vanessa Minden,
    Ülo Niinemets,
    Yusuke Onoda,
    Enio Egon Sosinski,
    Nadejda A. Soudzilovskaia,
    Michael C. Dietze
output: pdf_document
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage[table]{xcolor}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
---

```{r libraries, include = FALSE}
library(knitr)
library(kableExtra)
library(drake)
```

```{r sample_size_table, echo = FALSE, results = "asis"}
kable(
  readd(sample_size_wide),
  col.names = c(
    "PFT", "Leaf lifespan", "SLA",
    "$N_\\textrm{mass}$", "$N_\\textrm{area}$",
    "$P_\\textrm{mass}$", "$P_\\textrm{area}$",
    "$R_\\textrm{d, mass}$", "$R_\\textrm{d, area}$",
    "$V_\\textrm{c, max, mass}$", "$V_\\textrm{c, max, area}$",
    "$J_\\textrm{max, mass}$", "$J_\\textrm{max, area}$"
  ),
  format = "latex",
  booktabs = TRUE,
  escape = FALSE,
  caption = "The number of non-missing (outside parentheses) and imputed (inside parentheses) observations of each trait for each plant functional type used in this analysis."
) %>%
  kable_styling(latex_options = c("scale_down", "hold_position")) %>%
  group_rows("Broadleaf", 1, 5) %>%
  group_rows("Needleleaf", 6, 8) %>%
  group_rows("Shrub", 9, 11) %>%
  group_rows("Grass", 12, 14)
```

```{r pairwise_correlation_table, echo = FALSE, results = "asis"}
kable(
  readd(corr_processed), "latex",
  booktabs = TRUE, longtable = TRUE, escape = FALSE,
  col.names = c("PFT", "Trait 1", "Trait 2", "Correlation (95\\% CI)"),
  caption = "Pairwise trait correlation values for each plant functional type. Values with asterisks indicate correlations significantly different from zero ($p < 0.05$)."
)
```
